%\documentclass[aps,prl,twocolumn,nobibnotes]{revtex4}
\documentclass[aps,showpacs,twocolumn,nobibnotes]{revtex4}
%\documentclass[aps,preprint,showpacs,nobibnotes]{revtex4}
%\documentclass[aps,preprint,nobibnotes]{revtex}
\usepackage{graphics,graphicx,amsfonts,amsmath,amsbsy,amssymb,color}
\usepackage{bm}
%\usepackage{epic}
%\usepackage{mciteplus}
\usepackage{subfigure}
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\newcommand{\D}[1] {D_{\bf #1}}
\def \beq {\begin{eqnarray}}
\def \eeq {\end{eqnarray}}
\def \Schrodinger {{Schr\"{o}dinger }}
\def \Di {{D_{\bfi}}}
\def \Dj {{D_{\bfj}}}
\newcommand {\adag}[1] {{a_{#1}^\dagger}}
\def \bfj {{\bf j}}
\def \bfi {{\bf i}}
\def \rone {{\bvec{r}_1}}
\def \rtwo {{\bvec{r}_2}}
\def \mEh {{\textrm{mE}_{\textrm{h}}}}
\def \Eh {{\textrm{E}_{\textrm{h}}}}
\def \nadd {{n_a}}
\newcommand{\braket}[3] {{\langle #1 | #2 | #3 \rangle}}
\newcommand{\brket}[2] {{\langle #1 | #2 \rangle}}
\newcommand{\bra}{\ensuremath{\langle}}
\newcommand{\ket}{\ensuremath{\rangle}}
%\def \ham {{\bf H}}
\def \ham {{\hat{H}}}
%\def \Sz {{\hat{\textrm{S}_{\textrm{z}}}}}
\def \Sz {{\hat{S}_z}}
\newcommand{\rff}[1]{{Eq.~\eqref{#1}}}
\def \Pgen {{P_{\textrm{gen}}}}
\def \Carb {{\textrm{C}_{\textrm{2}}}}
\def \Hij {{H_{\bvec{i}\bvec{j}}}}
\def \Kii {{K_{\bvec{i}\bvec{i}}}}
\def \Kij {{K_{\bvec{i}\bvec{j}}}}

\begin{document}
\title{Spectral functions of extended systems via quantum embedding}
\author{George~H.~Booth}
%\email{ghb24@cam.ac.uk}
\author{Garnet~Kin-Lic~Chan}  
\affiliation{Department of Chemistry, Frick Laboratory, Princeton University, Princeton, New Jersey 08544, USA}

\begin{abstract}
In a previous publication [PRL {\bf 109} 186404 (2012)], the ground-state density matrix embedding theory (DMET) was introduced. With similarities 
to the dynamical mean-field method, a set of local sites are self-consistently correlated with a bath describing the complete mean-field 
coupling to the rest of the extended system. Despite many formal advantages in the analytic embedding, the method was restricted to ground-state properties.
Here, we extend the approach to introduce a frequency dependence, and demonstrate accurate spectral functions at a small 
computational cost. In a similar spirit to the frequency independent DMET method, the impurities are coupled via a small set of 
algebraically constructed, but now frequency-dependent bath states. The resultant spectral functions are directly obtained on the 
real-frequency axis, and allow for straightforward generalization both to multiple impurities and arbitrary perturbation operators. We demonstrate
the application of this method on the hubbard model, for both quasiparticle and optical spectra. This vastly extends the scope and applicability 
of the DMET method in condensed matter problems as a cheap route to correlated local spectral functions of extended systems.
\end{abstract}
\date{\today}
\maketitle

Dynamic correlation functions are directly probed in spectroscopic techniques, and correspond to many of the important transport and 
wider electronic structure properties of materials. As such, their accurate theoretical construction is key for materials scientists. 
However, few robust approaches exist for strongly correlated problems. The difficulty is in simultaneously requiring both an accurate 
treatment of the electron correlations beyond mean-field band theory for all excited state components at each frequency, as well as 
convergence with respect to the finite-size effects in the solid state. In dynamical mean-field theory (DMFT), the central quantity 
is the greens function of the system, which is self-consistenty optimized 

Size of bath independent of the size of the underlying lattice. Only a function of the impurity cluster size and the perturbation, while still exact in the non-interacting limits.

Schmidt decompose the action of the non-interacting operator, into that acting on the space of the impurity+bath ground state system, and that acting on the rest of the environment orbitals.

Cost of any frequency point not more than a minute on a single processor core.

%\bibliography{SpectralDMETBib}

\end{document}
